---
title: "GNG Analysis: open science"
author: "Mindy Ross & Theresa Nguyen"
date: "Updated `r Sys.Date()`"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Examing the effect of time of day on go/no-go task performance in those with mood disorders

```{r}
# import libraries
library(readr)
library(tidyverse)
library(tableone)
library(lmerTest)
library(sjPlot)
library(effects)
library(ggpubr)

# read in GNG file
f <- '/home/mindy/Desktop/BiAffect-iOS/data_processing/processed_outputs/GNG/AllUsers_GNGdata_fromZIP_allResponseTimes.csv'
dfRaw <- read.csv(f)
dfRaw$hour <-as.numeric(lubridate::hour(dfRaw$sessionTimestampLocal))
dfRaw$hour <- lubridate::hour(dfRaw$sessionTimestampLocal)
# remove rxn times > 3 sec
dfRaw$timeToThreshold_all <- ifelse(dfRaw$timeToThreshold_all > 3, NA, dfRaw$timeToThreshold_all)

dfRaw$genderFactor <- ifelse(dfRaw$gender == 'Male', 0,
                             ifelse(dfRaw$gender == 'Female', 1,
                                    ifelse(dfRaw$gender == 'Non-binary or third gender', 2, NA)))
dfRaw <- dfRaw %>% filter(genderFactor != 2)
dfRaw$genderFactor <- factor(dfRaw$genderFactor, levels=c(0,1),labels=c('Male','Female'))

dfRaw$diagnosis <- replace(dfRaw$diagnosis, dfRaw$diagnosis == '', NA) 
dfRaw <- dfRaw %>% drop_na(diagnosis)
dfRaw$diagnosis <- factor(dfRaw$diagnosis) 

dfRaw$BD_factor <- factor(dfRaw$BD, levels=c(0,1),labels=c('Negative','Positive'))
dfRaw$depression_factor <- factor(dfRaw$depression, levels=c(0,1),labels=c('Negative','Positive'))
dfRaw$SAD_factor <- factor(dfRaw$SAD, levels=c(0,1),labels=c('Negative','Positive'))
dfRaw$anxiety_factor <- factor(dfRaw$anxiety, levels=c(0,1),labels=c('Negative','Positive'))
dfRaw$ADHD_factor <- factor(dfRaw$ADHD, levels=c(0,1),labels=c('Negative','Positive'))
dfRaw$schizophrenia_factor <- factor(dfRaw$schizophrenia, levels=c(0,1),labels=c('Negative','Positive'))
dfRaw$OCD_factor <- factor(dfRaw$OCD, levels=c(0,1),labels=c('Negative','Positive'))
dfRaw$PTSD_factor <- factor(dfRaw$PTSD, levels=c(0,1),labels=c('Negative','Positive'))
dfRaw$SUD_factor <- factor(dfRaw$SUD, levels=c(0,1),labels=c('Negative','Positive'))

# add 1 sec as the correct no-go rxn time
dfRaw$rxnTimeToPlot <- with(dfRaw, ifelse((go == 'False') & (incorrect == 'False'), 1, timeToThreshold_all))

set.seed(1234)
dfRxn <- dfRaw %>% filter((timeToThreshold_all >= 0.2) & (timeToThreshold_all <= 1) & (hour > 7)) %>% 
                mutate(rxnTime =log(timeToThreshold_all),
                     rxnTime_shuffled = sample(rxnTime),
                     trial_type = factor(ifelse((go=='True'), 0,1),
                                  levels = c(0,1), labels = c('Go Trial','No-Go Trial'))) %>% ungroup() %>%
                filter((taskNumber > 1) & (taskNumber <= 21)) %>%
                group_by(userID) %>% mutate(nTasks = max(taskNumber)) %>% ungroup() %>%
                filter(nTasks >= 3) %>% drop_na(age,gender, diagnosis)

dfAccuracy <- dfRaw %>% filter((go == 'False') & (hour > 7)) %>% 
                mutate(nogo_flag = factor(ifelse(incorrect == 'False',1,0),levels=c(0,1),labels=c('Incorrect','Correct')), #incorrect=0, correct=1
                       nogo_flag_shuffled = sample(nogo_flag)) %>% ungroup() %>%
                filter((taskNumber > 1) & (taskNumber <= 21)) %>%
                group_by(userID) %>% mutate(nTasks = max(taskNumber)) %>% ungroup() %>%
                filter(nTasks >= 3) %>% drop_na(age,gender, diagnosis)

unmasckUserList <- c("9bn-MeLyPNvGOsB3GTgX7lfF", "nFo64ndW6U-RVbNL5IEd9ck6", "XKczkbtmSoX3pv4mFLkbDcTA","DUcdX56Np_058YZm1mmGXfZu", "E2--i7mqXTdtUrvvE0G-TTB6", "G5A91eHtAnaA4PIKWszot9Cn", "iSX3--I3H49YxnxMYcuIc8Ok", "bECjQIc1NEs1MdU84SPIsVzR", "yuRtbuOXxVYU3DlkrbRTbQxm","JBxb6dnWv9sA06VhyeaIaZgf", "jodDi7Lxp0fMGBug74uqoyX8", "6-VjFj3_X1ciuGCyGC_q1-_i","iCHdqkYDvNqHhzUHWYZmdfq7", "5GtX_dZueSvqJCLEvk2Fnl9q", "CYkG7LTpJ0k8UFai6aHOgKeo","Ie1WNvTKJ2hjXfEF1gT9WSw7", "1E7M7HvzqaRy59yX1MPCNPDW", "DGzqEqovjkrGllxCkpcrsDTn","EVtPtSokvtvuvxK9msXVA4TM", "UOVJXxoxljfah4G_EF_oZLr1", "KpYwG_qGGtiR2nme6XENvyEj","Vofg7z1xADgFgFC7k13OHwSK", "8wK-7SqPRerEi1JCJE369DBz", "y2OxoEXgZQE-Z17g5u394Jbk","zzP_LG5FJtLSj4gCWjZ4roLJ", "kiPdi044EmtNAo8a5hdzQLgm", "0_POeg9ljwiSjEKfvb590BD_","ownnjCEVtv7TPmSfrlcgxz06", "jJ1whkA0UjUt0rQPsDkyhtEr", "zcY6E3SZoS58KTEkwEan_Mf3","7PvMXRMo_xaZJE-LWAr_efDk", "oprPi_AoMi6T-kv-cISRVQPK", "RsRtg9O1FhpbLYk38h3SPWAw","fmELsbWMN1s3rZKKfHKOA5Ei", "NZnq9pGcBex8jx61KYV8g5nF", "GMhEgDi7Z9GiHLQVA9SacL77", "FBJNri-kuLEc6yzJDOly8Xh-", "ywJFPniW7HI7dyQUOJIED0F7", "i7nrQnZ_phP5UJnEHpdHgZ8L", "1jOrjCim1k38slSSP3xLdeWS", "SCeDzbFG67htYpuY__6kp2te", "HHL0VrVtuTR76ehtmdOj406L", "hoSZBt59vCMGzZMbcl7q0Ms9", "BLgOI8rHmAp2452xnQHHBt7b", "LjD0SMwPGDoPq4XQ-bAaMMH9","wZc40txUbq6BCDbHyU9bVcJ9", "pwJtx0eaiOcSVQ7yycsfWdLR", "K-HgXY6Ee9-g-zI9JUZAVYXI", "8NPIpWwE2HtW9Ms6MVr4iB8D", "2Frbad2afSuc8AyfoprxyBE-", "PW0HPDKMrE1sUYgJe3y1GW0z","ArrRKDCDA68U951hyrpmGXrl", "1p-U7v3p8UWcwHAxpRk7xS6R", "bFzc7ZZiZw0TYHRP-H8__qUM","ZqQCbaijbsz4zPioRh9TO58g", "47rZmq7WzvkFN1Xuz3iz7M5L", "lRGPI3JBxSqvss_cYmyT0gS8","dl3zJf2WXYMY36hZZ3FBzIhH", "aBmWuugUK_0Kc6dScUkg1BLi", "l0DjOCoJ-s9OfNTEx90ptDlH","x35Fz4DEhyvEScX7p2JuZu4T", "ilDvKT9fCZhlvuET0CG5eHNu", "A1regU9f5KHP9RsyTDSnw-8e", "ZUsw-e-L5xL_PYVrjEdHzYzc", "PpVWhdzXE_2YvvwxtgTMkX-0", "pmUSIo7ZoyIcPm7kFxYtMwQN","6LLSVAHYT2RAJj0jHOz832S4", "McJIiBBJo8a8g7i_d1pgSojE", "AFLmMCDnHqiVawlfl9GUO4pe","XAy4Wa1IrUIKXSrLkZb5eKpy", "30NTdwLBHz1z8RYCECUhnFrO")
OSuserList <- unique(dfRxn$healthCode)

table(unmasckUserList %in% OSuserList)
print('two common users - remove both')

# remove common users
dfRxn <- dfRxn %>% filter(!(healthCode %in% unmasckUserList))
dfAccuracy <- dfAccuracy %>% filter(!(healthCode %in% unmasckUserList))

dfDemo <- dfRxn %>% group_by(userID) %>% 
                summarise(gender = unique(genderFactor),
                          age = unique(age),
                          diagnosis = unique(diagnosis),
                          BD = unique(BD_factor),
                          depression = unique(depression_factor),
                          SAD = unique(SAD_factor),
                          anx = unique(anxiety_factor),
                          ADHD = unique(ADHD_factor),
                          schizo = unique(schizophrenia_factor),
                          ocd = unique(OCD_factor),
                          ptsd = unique(PTSD_factor),
                          sud = unique(SUD_factor),
                          nGNG = max(taskNumber)) 
```

## Demographics Summary

```{r}
tab3 <- CreateTableOne(data = subset(dfDemo, select= -c(userID, diagnosis)))
ptables <- print(tab3,  quote = FALSE, noSpaces = TRUE, printToggle = FALSE, test = TRUE, showAllLevels = FALSE, missing = FALSE, varLabels = TRUE )
ptables
```

## Roster Plot

```{r}
dfRxn$date <- as.Date(dfRxn$date, "%Y-%m-%d")
dfRxn2 <- dfRxn %>% group_by(userID) %>% mutate(dayNumber = as.integer(date - min(date) + 1)) %>% ungroup %>%
                    group_by(userID, taskNumber) %>% filter(row_number()==1)

ggplot(dfRxn2, aes(x=dayNumber, y = userID, group = taskNumber)) + 
  geom_count()

dfRxn3 <- subset(dfRxn2, dfRxn2$dayNumber <50)
ggplot(dfRxn3, aes(x=dayNumber, y = userID, group = taskNumber)) + 
  geom_count()

dfRxn4 <- subset(dfRxn2, dfRxn2$dayNumber <30)
ggplot(dfRxn4, aes(x=dayNumber, y = userID, group = taskNumber)) + 
  geom_count()
```

## Age Distribution

```{r}
ageList <- dfRxn %>% group_by(userID) %>% summarise(age = unique(age))
hist(ageList$age, breaks=20)
```

## GNG distributions

**Hourly task distribution**

GNG tasks tend to be completed between the hours of 8-23 (0800-2359). Data filtered in model to only include tasks completed during these hours.

```{r}
hist(dfRxn$hour, breaks=35)
```

**GNG Task Number Distribution**

```{r}
tasks <- dfRaw %>% group_by(userID) %>% summarise(n = unique(taskNumber))
h<-hist(tasks$n, breaks=35)

# median # tasks per person
median(tasks$n)
```

**Total number of GNG tasks per person in models**

```{r}
modelTasks <- dfRxn %>% group_by(userID) %>% summarise(max = max(taskNumber))
hist(modelTasks$max)
```

**GNG Accuracy Distribution**

Fraction of correct responses to no-go trials

```{r}
dfAccPerc <- dfAccuracy %>% group_by(userID, taskNumber) %>% summarise(percAccurate = mean(ifelse(incorrect == 'False',1,0)))
hist(dfAccPerc$percAccurate)
```

Histogram of no-go trial accuracy 

* 1 is correct
* 0 is incorrect

```{r}
hist(ifelse(dfAccuracy$incorrect == 'False',1,0))
```

# Mixed Models Analyses

## No-Go Trial Accuracy

### Effect of hour, diagnosis, and practice on no-go trial accuracy

Random intercept per subject

```{r}
model1Nogo <- glmer(nogo_flag ~ scale(age) + genderFactor +
                      scale(log(taskNumber)) + 
                      BD_factor +
                      depression_factor +
                      SAD_factor +
                      anxiety_factor +
                      ADHD_factor +
                      PTSD_factor +
                      poly(hour,2) +
                  (1 + scale(log(taskNumber)) | userID), 
                  data = dfAccuracy, family = 'binomial',
                  control=glmerControl(optimizer="bobyqa"))
summary(model1Nogo)
car::vif(model1Nogo)
```

### Model Summaries

```{r}
tab_model(model1Nogo,digits=3, digits.re = 3,show.ci=FALSE, show.loglik = TRUE, show.obs = TRUE, show.ngroups = TRUE, show.icc=TRUE)
```

### Plots

```{r}
effectNumNogo <- as.data.frame(effects::Effect('taskNumber', model1Nogo))
model1_plot_numNogo <- ggplot(data=effectNumNogo, aes(x=taskNumber, y=fit))+
  geom_line(size=1)+
  geom_ribbon(aes(ymin=lower, ymax=upper),alpha=.2) +
  xlab("Task Number")+
  ylab("No-Go Trial \nAccuracy")+
  theme_bw()+
  theme(text=element_text(face="bold", size=12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = "NA"),
        axis.line = element_line(size = 1, colour = "grey80"),
        legend.title=element_blank())

effectDep <- as.data.frame(Effect(c('depression_factor'), model1Nogo))
model1_plot_Dep <- ggplot(data=effectDep, aes(x=depression_factor, y=fit))+
  geom_point(size=4)+
  geom_errorbar(aes(ymin=lower, ymax=upper),alpha=.5, size=1) +
  xlab("Depression")+
  ylab("No-Go Trial \nAccuracy")+
  theme_bw()+
  theme(text=element_text(face="bold", size=12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = "NA"),
        axis.line = element_line(size = 1, colour = "grey80"),
        legend.title=element_blank())


figure1 <- ggarrange(model1_plot_numNogo,
                    model1_plot_Dep,
                     labels = c("A","B"), 
                     ncol = 2, nrow=1, widths=c(0.9,1))
ggsave(filename = 'nogo_OSData_300dpi.png', path = '/home/mindy/Desktop/BiAffect-iOS/UnMASCK/gng_rxn_time/paper/figures/', width = 9, height = 4, device='png', dpi=300)

```

## Reaction Time

### Effect of hour and trial type on reaction time

```{r}
model1Rxn <- lmer(rxnTime ~ scale(age) +
                    scale(log(taskNumber)) +  genderFactor + trial_type + 
                      BD_factor +
                      depression_factor +
                      SAD_factor +
                      anxiety_factor +
                      ADHD_factor +
                      PTSD_factor +
                    poly(hour,2) + 
                 (1 + scale(log(taskNumber)) | userID), data = dfRxn, REML=FALSE, control = lmerControl(optimizer='nmkbw'))
summary(model1Rxn)
car::vif(model1Rxn)
```

### Model Summaries

```{r}
tab_model(model1Rxn,digits=3, digits.re = 3,p.val = 'satterthwaite',show.ci=FALSE, show.loglik = TRUE, show.obs = TRUE, show.ngroups = TRUE, show.icc=TRUE)
```

### Plots

```{r}
effectAgeRxn <- as.data.frame(Effect(c('age'), model1Rxn))
model1_plot_ageRxn <- ggplot(data=effectAgeRxn, aes(x=age, y=exp(fit)))+
  geom_line(size=1)+
  geom_ribbon(aes(ymin=exp(lower), ymax=exp(upper)),alpha=.2) +
  xlab("Age")+
  ylab("GNG Reaction \nTime (s)")+
  theme_bw()+
  theme(text=element_text(face="bold", size=12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = "NA"),
        axis.line = element_line(size = 1, colour = "grey80"),
        legend.title=element_blank())

effectHourRxn <- as.data.frame(Effect(c('hour'), model1Rxn))
model1_plot_hrRxn <- ggplot(data=effectHourRxn, aes(x=hour, y=exp(fit)))+
  geom_line(size=1)+
  geom_ribbon(aes(ymin=exp(lower), ymax=exp(upper)),alpha=.2) +
  xlab("Hour")+
  ylab("GNG Reaction \nTime (s)")+
  theme_bw()+
  theme(text=element_text(face="bold", size=12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = "NA"),
        axis.line = element_line(size = 1, colour = "grey80"),
        legend.title=element_blank())

effectTrialRxn <- as.data.frame(Effect(c('trial_type'), model1Rxn))
model1_plot_trialRxn <- ggplot(data=effectTrialRxn, aes(x=trial_type, y=exp(fit)))+
  geom_point(size=4)+
  geom_errorbar(aes(ymin=exp(lower), ymax=exp(upper)),alpha=.5, size=1) +
  xlab("Trial Type")+
  ylab("GNG Reaction \nTime (s)")+
  theme_bw()+
  theme(text=element_text(face="bold", size=12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = "NA"),
        axis.line = element_line(size = 1, colour = "grey80"),
        legend.title=element_blank())

figure2 <- ggarrange(model1_plot_ageRxn,
                    model1_plot_trialRxn,
                    model1_plot_hrRxn,
                     labels = c("A","B","C"), 
                     ncol = 1, nrow=3, widths=c(0.9,1))
figure2
ggsave(filename = 'rxn_OSData_300dpi.png', path = '/home/mindy/Desktop/BiAffect-iOS/UnMASCK/gng_rxn_time/paper/figures/', width = 5, height = 10, device='png', dpi=300)
```

### Histogram of Reaction Times

```{r}
ggplot(data=dfRxn, aes(x=exp(rxnTime), group = trial_type))+
  geom_histogram(aes(fill=trial_type)) +
  xlab("Time (s)")+
  ylab("Count")+
  theme_bw()+
  theme(text=element_text(face="bold", size=12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = "NA"),
        axis.line = element_line(size = 1, colour = "grey80"),
        legend.title=element_blank())
```