---
title: "GNG Analysis: UnMASCK"
author: "Mindy Ross & Theresa Nguyen"
date: "Updated `r Sys.Date()`"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Examing the effect of time of day on go/no-go task performance in those with mood disorders

```{r}
# import libraries
library(readr)
library(tidyverse)
library(tableone)
library(lmerTest)
library(sjPlot)
library(effects)
library(ggpubr)

# read in GNG file
f <- '/home/mindy/Desktop/BiAffect-iOS/UnMASCK/BiAffect_data/processed_output/GNG/AllUsers_GNGdata_fromZIP_allResponseTimes_v3.csv'
dfRaw <- read.csv(f)
dfRaw$hour <-as.numeric(lubridate::hour(dfRaw$sessionTimestampLocal))
dfRaw$hour <- lubridate::hour(dfRaw$sessionTimestampLocal)
dfRaw$hour2 <- ifelse(dfRaw$hour == 0,24,dfRaw$hour)
dfRaw$gender <- dfRaw$gender-1
dfRaw$gender <- factor(dfRaw$gender, levels=c(0,1),labels=c('Male','Female'))
dfRaw$diagnosis <- dfRaw$diagnosis-1
dfRaw$diagnosis <- factor(dfRaw$diagnosis, levels=c(0,1), labels=c('HC','MD'))
dfRaw$diagnosis_group <- factor(dfRaw$diagnosis_group, levels=c(0,1,2), labels=c('HC','D','BD'))
# add 1 sec as the correct no-go rxn time
dfRaw$rxnTimeToPlot <- with(dfRaw, ifelse((go == 'False') & (incorrect == 'False'), 1, timeToThreshold_all))

set.seed(123)
dfPlot <- dfRaw %>% filter((rxnTimeToPlot >= 0.2) & (rxnTimeToPlot <= 1) & (hour2 > 6)) %>% 
              mutate(trial_type = factor(ifelse((go=='True')&(incorrect=='False'), 0, #correct go
                                         ifelse((go=='False')&(incorrect=='True'), 1, #incorrect no go
                                         ifelse((go=='False')&(incorrect=='False'), 2, NA))), # correct no go
                          levels = c(0,1,2), labels = c('correct_go_trial','incorrect_noGo_trial','correct_noGo_trial'))) %>% 
                          ungroup %>% group_by(userID) %>% mutate(hour_range = max(hour2) - min(hour2)) %>% 
                          filter((hour_range >= 8) & (max(taskNumber) >= 10))

dfRxn <- dfRaw %>% filter((timeToThreshold_all >= 0.2) & (timeToThreshold_all <= 1) & (hour2 > 6)) %>% 
                mutate(rxnTime =log(timeToThreshold_all),
                     rxnTime_shuffled = sample(rxnTime),
                     trial_type = factor(ifelse((go=='True'), 0,1),
                                  levels = c(0,1), labels = c('Go Trial','No-Go Trial'))) %>% 
                ungroup %>% group_by(userID) %>% mutate(hour_range = max(hour2) - min(hour2)) %>% 
                          filter((hour_range >= 8) & (max(taskNumber) >= 10)) %>% ungroup() %>%
                drop_na(gender, age, visit1_QIDS)

set.seed(1234)
dfAccuracy <- dfRaw %>% filter((go == 'False') & (hour2 > 6)) %>% 
                mutate(nogo_flag = factor(ifelse(incorrect == 'False',1,0),levels=c(0,1),labels=c('Incorrect','Correct')), #incorrect=0, correct=1
                       nogo_flag_shuffled = sample(nogo_flag)) %>%
              ungroup %>% group_by(userID) %>% mutate(hour_range = max(hour2) - min(hour2)) %>% 
                          filter((hour_range >= 8) & (max(taskNumber) >= 10)) %>% ungroup() %>% drop_na(gender, age, visit1_QIDS)

dfDemo <- dfRxn %>% group_by(userID) %>% 
                summarise(gender = unique(gender),
                          age = unique(age),
                          diagnosis = unique(diagnosis),
                          diagnosis_group = unique(diagnosis_group),
                          visit1_QIDS = unique(visit1_QIDS),
                          nGNG = max(taskNumber)) 
```

## Raster Plot

```{r}
dfRxn$date <- as.Date(dfRxn$date, "%Y-%m-%d")
dfRxn2 <- dfRxn %>% group_by(userID) %>% mutate(dayNumber = as.integer(date - min(date) + 1)) %>% ungroup %>%
                    group_by(userID, dayNumber,taskNumber) %>% filter(row_number()==1) %>% ungroup %>%
                    group_by(userID, dayNumber) %>% mutate(count = n())

ggplot(dfRxn2, aes(x=dayNumber, y = userID, group = count)) + 
  geom_point(size=count)

dfRxn3 <- subset(dfRxn2, dfRxn2$dayNumber <40)
ggplot(dfRxn3, aes(x=dayNumber, y = userID, group = count)) + 
  geom_point(aes(size=count))
```

```{r}
ggplot(data = dfRxn, aes(x=hour2, fill=diagnosis)) +
  geom_histogram(aes(color=diagnosis))
```

## Demographics Summary

```{r}
tab3 <- CreateTableOne(strata = "diagnosis", data = subset(dfDemo, select=-c(userID)))
ptables <- print(tab3,  quote = FALSE, noSpaces = TRUE, printToggle = FALSE, test = TRUE, showAllLevels = FALSE, missing = FALSE, varLabels = TRUE )
ptables
```

- Note: ignore row with diagnosis in table above

## QIDS distribution

```{r}
hist(dfRxn$visit1_QIDS)
```

## GNG distributions

**Hourly task distribution**

GNG tasks tend to be completed between the hours of 7-24 (0700-0059). Data filtered in model to only include tasks completed during these hours.

```{r}
hist(dfRaw$hour, breaks=25)
```

**GNG Reaction Time**

Before filtering

```{r}
dfRxnTime <- subset(dfRaw, (dfRaw$timeToThreshold_all >0) & (dfRaw$timeToThreshold_all <=3))

ggplot(dfRxnTime, aes(x=timeToThreshold_all)) + 
  geom_histogram(bins=50) +
  scale_x_continuous(breaks=seq(0,3,0.2))+ 
  geom_vline(xintercept = 0.2,color='darkgrey') +
  geom_vline(xintercept = 1,color='darkgrey') +
  xlab("Reaction Time (s)")+
  ylab("Count")+
  theme_bw()+
  theme(text=element_text(face="bold", size=12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = "NA"),
        axis.line = element_line(size = 1, colour = "grey80"),
        legend.title=element_blank())

ggsave(filename = 'histogram_rxnTimes_300dpi.png', path = '/home/mindy/Desktop/BiAffect-iOS/UnMASCK/gng_rxn_time/paper/figures/', width = 6, height = 4, device='png', dpi=300)
```

All GNG reaction times (correct and incorrect trials) following filtering to between 0.2 - 1 sec.

```{r}
hist(exp(dfRxn$rxnTime))
```

Filtered logged GNG reaction time per trial (correct and incorrect trials)

```{r}
hist(dfRxn$rxnTime)
```

**Comparison of reaction time by trial type**

```{r}
ggplot(dfRxn, aes(x=exp(rxnTime), group = trial_type)) + 
  geom_boxplot(aes(fill=trial_type)) +
  xlab("Reaction Time")+
  ylab("Trial Type")+
  theme_bw()+
  theme(text=element_text(face="bold", size=12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = "NA"),
        axis.line = element_line(size = 1, colour = "grey80"),
        legend.title=element_blank())
```

## Spaghetti Plots

Correct no-go trials given a reaction time of 1 sec in the plots below to visualize their frequency relative to the other trials with reaction times. Incorrect go trials were not plotted.

```{r,fig.width=12,fig.height=12}
ggplot(dfPlot, aes(x=jitter(hour2), y=jitter(rxnTimeToPlot), trial_type)) + 
      geom_point(size=0.25, aes(color=trial_type, shape=trial_type), alpha=0.7) + 
      scale_color_manual(values=c('#56B4E9','red','chartreuse4')) +
      scale_shape_manual(values=c(3,19,3)) +
      facet_wrap(~userID)
```

## No-go trial counts by hour and trial accuracy

Incorrect = False means correct trial
Incorrect = True means incorrect trial

# Mixed Models Analyses

## No-Go Trial Accuracy

### Effect of hour, subject, and practice on no-go trial accuracy

Random slope of hour and intercept per subject

```{r}
model1Nogo <- glmer(nogo_flag ~ scale(age) + gender + 
                      scale(log(taskNumber)) +  scale(visit1_QIDS) * poly(hour2,2, raw=FALSE) +    
                  (1 + scale(log(taskNumber)) | userID), 
                  data = dfAccuracy, family = 'binomial',
                  control=glmerControl(optimizer="bobyqa")) 
summary(model1Nogo)
car::vif(model1Nogo)
```

### Model Summaries

```{r}
tab_model(model1Nogo,digits=3, digits.re = 3,show.ci=FALSE, show.loglik = TRUE, show.obs = TRUE, show.ngroups = TRUE, show.icc=TRUE)
```

### Plots

```{r}
effectNumNogo <- as.data.frame(effects::Effect('taskNumber', model1Nogo))
model1_plot_numNogo <- ggplot(data=effectNumNogo, aes(x=taskNumber, y=fit))+
  geom_line(size=1)+
  geom_ribbon(aes(ymin=lower, ymax=upper),alpha=.2) +
  xlab("Task Number")+
  ylab("No-Go Trial \nAccuracy")+
  theme_bw()+
  theme(text=element_text(face="bold", size=12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = "NA"),
        axis.line = element_line(size = 1, colour = "grey80"),
        legend.title=element_blank())

intPred_nogo <- ggeffects::ggpredict(model1Nogo, terms = c("hour2 [all]", "visit1_QIDS [0,5,10]"))
intPlot_nogo <- ggplot(intPred_nogo, aes(x = x, y = predicted, colour = group, fill=group)) +
  geom_line(size=1) +
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high), alpha=.1,  colour = "NA") +
  xlab("Hour")+
  ylab("No-Go Trial \nAccuracy") +
  theme_bw() +
  theme(text=element_text(face="bold", size=12),
        panel.border = element_rect(fill = NA, colour = "NA"),
        axis.line = element_line(size = 1, colour = "grey80"),
        legend.position = 'top',
        legend.title=element_blank())

figure1 <- ggarrange(model1_plot_numNogo,
                    intPlot_nogo,
                     labels = c("A","B"), 
                     ncol = 2, nrow=1, widths=c(0.9,1))
figure1

ggsave(filename = 'nogo_unmasckData_v2.png', path = '/home/mindy/Desktop/BiAffect-iOS/UnMASCK/gng_rxn_time/paper/figures/', width = 10, height = 4.5, device='png', dpi=300)
```

## Reaction Time

### Effect of hour and trial type on reaction time

```{r}
model1Rxn <- lmer(rxnTime ~ scale(age) + gender + 
                    scale(log(taskNumber)) + trial_type + scale(visit1_QIDS) * poly(hour2,2, raw=FALSE) + 
                 (1 + scale(log(taskNumber)) | userID), data = dfRxn, REML=FALSE, control = lmerControl(optimizer='nmkbw'))
summary(model1Rxn)
car::vif(model1Rxn)
```

### Model Summaries

```{r}
tab_model(model1Rxn,digits=3, digits.re = 3,p.val = 'satterthwaite',show.ci=FALSE, show.loglik = TRUE, show.obs = TRUE, show.ngroups = TRUE, show.icc=TRUE)
```

### Plots

```{r} 
effectTrialRxn <- as.data.frame(Effect(c('trial_type'), model1Rxn))
model1_plot_trialRxn <- ggplot(data=effectTrialRxn, aes(x=trial_type, y=exp(fit)))+
  geom_point(size = 4)+
  geom_errorbar(aes(ymin=exp(lower), ymax=exp(upper)),alpha=.5, size = 1) +
  xlab("Trial Type")+
  ylab("GNG Reaction \nTime (s)")+
  theme_bw()+
  theme(text=element_text(face="bold", size=12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = "NA"),
        axis.line = element_line(size = 1, colour = "grey80"),
        legend.title=element_blank())

effectTrialRxn <- as.data.frame(Effect(c('trial_type'), model1Rxn))
model1_plot_trialRxn_withRaw <- ggplot()+
  geom_point(data=effectTrialRxn, aes(x=trial_type, y=exp(fit)), size = 4)+
  geom_errorbar(data=effectTrialRxn, aes(x=trial_type, ymin=exp(lower), ymax=exp(upper)),alpha=.5, size = 1) +
  geom_jitter(data=dfRxn, aes(x=trial_type, y=exp(rxnTime)), alpha=0.2, size=0.2)+
  xlab("Trial Type")+
  ylab("GNG Reaction \nTime (s)")+
  theme_bw()+
  theme(text=element_text(face="bold", size=12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = "NA"),
        axis.line = element_line(size = 1, colour = "grey80"),
        legend.title=element_blank())
model1_plot_trialRxn_withRaw

intPred <- ggeffects::ggpredict(model1Rxn, terms = c("hour2 [all]", "visit1_QIDS [0,5,10]"))
intPlot_rxn <- ggplot(intPred, aes(x = x, y = exp(predicted), colour = group, fill=group)) +
  geom_line(size=1) +
  geom_ribbon(aes(ymin=exp(conf.low), ymax=exp(conf.high)), alpha=.1,  colour = "NA") +
  xlab("Hour")+
  ylab("GNG Reaction \nTime (s)") +
  theme_bw() +
  theme(text=element_text(face="bold", size=12),
        panel.border = element_rect(fill = NA, colour = "NA"),
        axis.line = element_line(size = 1, colour = "grey80"),
        legend.position = 'right',
        legend.title=element_blank())

dfRxn$QIDS_grouped <- ifelse(dfRxn$visit1_QIDS < 5, 0,
                             ifelse(dfRxn$visit1_QIDS <10, 5,10))
intPlot_rxn_withRaw <- ggplot() +
  geom_line(data = intPred, aes(x = x, y = exp(predicted), colour = group, fill=group), size=1) +
  geom_ribbon(data = intPred, aes(x = x, y = exp(predicted), ymin=exp(conf.low), ymax=exp(conf.high), fill = group), alpha=.1) +
  geom_point(data = dfRxn, aes(x=hour2, y=exp(rxnTime), color=factor(QIDS_grouped)), alpha=0.1) +
  xlab("Hour")+
  ylab("GNG Reaction \nTime (s)") +
  theme_bw() +
  theme(text=element_text(face="bold", size=12),
        panel.border = element_rect(fill = NA, colour = "NA"),
        axis.line = element_line(size = 1, colour = "grey80"),
        legend.position = 'right',
        legend.title=element_blank())
intPlot_rxn_withRaw

figure2 <- ggpubr::ggarrange(model1_plot_trialRxn_withRaw, 
                             intPlot_rxn_withRaw,
                     labels = c("A","B"), 
                     ncol = 2, nrow=1, widths=c(0.9,1))
figure2

ggsave(filename = 'rxn_unmasckData_v2.png', path = '/home/mindy/Desktop/BiAffect-iOS/UnMASCK/gng_rxn_time/paper/figures/', width = 10, height = 4, device='png', dpi=300)
```